\documentclass[a4paper,12pt]{article}

\usepackage[pdftex]{graphicx}
\usepackage{amsmath,amsfonts}

\title{Test 3}
\date{Time: 4 hours}
\author{April Camp 2017}

\begin{document} \maketitle

\begin{enumerate}
	\item %PAMO 2009 Shortlist, A2 
	Given $5$ positive real numbers, show that there exist two of them,
	$a$ and $b$, for which
	\[
		0 \leq \frac{a}{1+a^2} - \frac{b}{1+b^2} \leq \frac{1}{8}.
	\]

	\item %PAMO 2009 Shortlist, G2
	An acute-angled triangle $ABC$ is given. Points $D$, $E$, and $F$ lie
	on the sides $BC$, $CA$, and $AB$ respectively, and satisfy $\angle
	FDE = \angle BAC$. and $\angle DEF = \angle ABC$. Prove that the
	orthocenter of triangle $DEF$ coincides with the circumcenter of
	triangle $ABC$.

	\item 
	Prove that for every prime $p$, there exists an integer $x$ such that \[x^8 \equiv 16 \pmod p.\]

	\item % MEMO 2011
	Let $n > 3$ be an integer. John and Mary play the following game: First John labels the sides of a regular $n$-gon with the numbers $1, 2, \dots, n$ in whatever order he wants, using each number exactly once. Then Mary divides this $n$-gon into triangles by drawing $n - 3$ diagonals which do not intersect each other inside the $n$-gon. All these diagonals are labeled with number $1$. Into each of the triangles the product of the numbers on its sides is written. Let $S$ be the sum of those $n - 2$ products.
	
Determine the value of $S$ if Mary wants the number $S$ to be as small as possible and John wants $S$ to be as large as possible and if they both make the best possible choices.

\end{enumerate}

\medskip 

\hfill \emph{Each problem is worth 7 points.}
	
\end{document}
